---
title: "Supplementary information"
format: pdf
---

# Clustering Robustness checks

To assess the validity of our clustering into three classes, we performed measurements both on Silhouette scores and Dunn's index. 
Both tests confirm that the partition is the best one of those available.
In the case of the silhouette scores, the 3 classes solution returns the highest score (.31). 
Regarding the Dunn's index, some solutions have higher scores, but none within a small number of partitions (the 3 class solution outperforms those between 2 and 7 classes).

## Silhouette scores

```{r echo=FALSE, message= FALSE}

#install.packages("ggdendro")
#install.packages("tidytable")
#install.packages("ggtext")
#install.packages("cluster")
#install.packages("clValid")
#install.packages("patchwork")
library(cluster)
library(clValid)

conversion_tables <- read_rds("conversion_table.RDS")
disc_clust_hc <- read_rds("disc_clust_hc.RDS")
disc_top_profiles_clr <- read_rds("disc_top_profiles_clr.RDS")
pca_disc <- read_rds("pca_disc.RDS")

source("run_disc_clustering.R")
source("viz_disc_clustering.R")
source("abbrev_discs.R")

disc_clust_hc <-  run_disc_clustering(disc_top_profiles_clr)

#viz_disc_clustering(disc_clust_hc,conversion_tables)

library(cluster)

rownames(disc_top_profiles_clr) <- abbrev_discs(rownames(disc_top_profiles_clr))

dist_m <- dist(disc_top_profiles_clr)

for_silhouette <- silhouette(
  x = cutree(disc_clust_hc,3),
  dist_m
)

squared_dist <- dist_m * dist_m

sil <- tibble()

for (i in 2:10) {
  temp_sil <- silhouette(x = cutree(disc_clust_hc, k = i),
  squared_dist) |> as_tibble() |> 
    mutate(n_clust = i,
    n_clust = ifelse(str_length(n_clust) == 1, 
  paste0("0", n_clust),
as.character(n_clust)))

sil <- sil |> 
  bind_rows(temp_sil)

}

# 


```


```{r echo=FALSE}
ggplot(sil |> 
  arrange(n_clust, cluster, -sil_width) |> 
  group_by(n_clust) |> 
  mutate(av_sil = paste0("k = ", n_clust, "; sil = ",
                         round(mean(sil_width), 2)),
         x = row_number()),
 aes(x = x,
  y = sil_width,
  fill = factor(cluster)
  ))+
  geom_col()+
  facet_wrap(~av_sil, scales = "free_x")+
  theme_bw()+
  ggtitle("Silhouette score\n(2 to 9 classes solutions)")+
  theme(legend.position = "none")
```

## Dunn's index

```{r, echo= FALSE}

the_dunn <- tibble()
for (i in 2:20) {
  cluster <- cutree(disc_clust_hc,i)
  temp_dunn <- clValid::dunn(squared_dist, cluster) |> 
    enframe() |> 
    mutate(name = i)
  
the_dunn <- the_dunn |> 
  bind_rows(temp_dunn)

}

the_dunn <- the_dunn |> 
  rename(n_clust = 1, dunn_score = 2)

```


```{r, echo = FALSE}

ggplot(the_dunn |>
  filter(n_clust <= 10) |>
  mutate(col = ifelse(n_clust == 3, TRUE, FALSE)), aes(x = n_clust, y = dunn_score, fill = col))+
  geom_col()+
  theme_bw()+
  theme(legend.position = "none")+
  ggtitle("Dunn's index\n(2 to 9 classes solutions)")


```

